---
title: "01 Preparing the data"
date: "`r Sys.time()`"
author: Sebastian Sauer
params:
  dataset: small
output: 
  html_document:
    toc: TRUE
    number_sections: TRUE
---


# Setup


```{r global-knitr-options, include=FALSE}

knitr::opts_chunk$set(
  fig.pos = 'H',
  fig.asp = 0.618,
  fig.align='center',
  fig.width = 5,
  out.width = "100%",
  fig.cap = "", 
  fig.path = "chunk-img/",
  dpi = 300,
  # tidy = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.show = "hold")

#knitr::opts_knit$set(root.dir = normalizePath(".."))
```




```{r}
library("tidyverse")
library("here")  # paths made easy
library("assertthat")  # testing
library("VIM")  # missing values
library("janitor")  # data cleansing
library("futile.logger")  # logging
library("labelled")  #  excluding nagging attributes of columns
library("visdat")  # csv fingerprint
library("gt")  # html tables
```


```{r}
source(here("src", "01-constants.R"))
source(here("src", "funs", "prepare-data.R"))
source(here("src", "funs", "helper-funs.R"))
source(here("src", "funs", "prop_NA.R"))
```


# Init logger

```{r init-logger}
flog.appender(appender.file("covid-icu.log"))
flog.threshold(INFO)
flog.info("STARTING script 01")
flog.info("Logger inited")
```



# Prepare data




**Dataset used: `r dataset`**

## Import and prepare data

```{r prepare-data, results = "hide", message = FALSE, warning = FALSE}
d1 <- prepare_data(data_raw_path = data_raw_path, 
                   dv = "verlauf",
                   id = "fallnummer")
flog.info("prepare_data was run")
```


Dimensions of the data set:

```{r}
dim(d1)
```




# d1 fingerprint


```{r fingerprint-csv}
vis_dat(d1)
```




>   Each cell is colored according to its type. Yellow for strings, blue for integers, purple for decimals and gray for empty values.


# Remove ROWS with many missing

Are there rows with many missings?


```{r}
d1 %>% 
  rowwise() %>% 
  mutate(isna_sum =  sum_isna(c_across(everything()))) %>% 
  mutate(isna_prop = isna_sum / ncol(.))
```



```{r count-na-rowwise}
na_rowwise <-
  d1 %>% 
  rowwise() %>% 
  mutate(isna_sum =  sum_isna(c_across(everything()))) %>% 
  mutate(isna_prop = isna_sum / ncol(.)) %>% 
  ungroup()
```


```{r}
na_rowwise %>% 
  select(id, isna_sum, isna_prop) %>% 
  arrange(-isna_prop) %>% 
  filter(isna_prop > .3)
```





# Missing data analysis


```{r}
VIM::aggr(d1)
```


```{r comp-na}
d1_sumisna <- 
  d1 %>% 
  summarise(across(everything(), sum_isna)) %>% 
  pivot_longer(everything()) %>% 
  mutate(prop = value/nrow(d1)) %>% 
  arrange(-prop) %>% 
  filter(prop > .1)

flog.info("Vars with more than 10% NA to be excluded")
```


# Remove variables with more than 10% missings

Which variables have more than 10% missing values?

```{r show-prop-na}
prop_NA(d1) %>% 
  gt() %>% 
  fmt_number(3)
```


What are the names of those variables (With many missings)?

```{r show-cols-many-missings}
vars_many_missings <- 
  d1 %>% 
  prop_NA %>% 
  pull(name)

vars_many_missings
```



# d2: data without vars with many missings


```{r rm-cols-with-many-missings}
d2 <-
  d1 %>% 
  select(-all_of(vars_many_missings))

flog.info("Remove vars with many missings (>10%)")
```




# Recode dependent variable

```{r recode-dv}
d2 <- 
  d2 %>% 
  mutate(verlauf = as.factor(verlauf)) %>% 
  mutate(verlauf = case_when(
    verlauf == 1 ~ "good",
    verlauf == 2 ~ "bad"
  )) 

flog.info("Recoded DV to 'good', 'bad'")
```

```{r}
dim(d2)
```


# Exclude (near) zero variance (nzv) variables


```{r find-nzv}
nzv_vars <- caret::nearZeroVar(d2, names = TRUE)
nzv_vars
```

```{r rm-nzv-vars}
d3 <-
  d2 %>% 
  select(-all_of(nzv_vars))

flog.info("Removed near zero variance variables")
```


# Excluding attributes


```{r remove-attirbues}
d3 <- labelled::remove_attributes(d3, "spec")

flog.info("Removing attributes 'spec' from df")
```


# Remaining columns in data set

```{r}
names(d3)
```




# Writing to disk

```{r write-to-disk}
write_csv(d3, file = data_processed_path)

flog.info(paste0("Written 'data-cleaned.csv' to disk to this path: ", data_processed_path))
```

