---
title: "02 descriptive analysis"
date: "`r Sys.time()`"
author: Sebastian Sauer
params:
  dataset: "small"
output: 
  html_document:
    number_sections: TRUE
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

# Setup


```{r global-knitr-options, include=FALSE}

knitr::opts_chunk$set(
  fig.pos = 'H',
  fig.asp = 0.618,
  fig.align='center',
  fig.width = 5,
  out.width = "100%",
  fig.cap = "", 
  fig.path = "chunk-img/",
  dpi = 300,
  # tidy = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.show = "asis")

#knitr::opts_knit$set(root.dir = normalizePath(".."))
```



```{r load-libs}
library("tidyverse")
library("here")
library("assertthat")
library("VIM")
library("janitor")
library("ggstatsplot")
library("glue")
library("gt")
library("rstatix")
library("broom")
library("futile.logger")
library("gt")
library("infer")
library("effectsize")

conflicted::conflict_prefer("cohens_d", "effectsize")
conflicted::conflict_prefer("discard", "purrr")
conflicted::conflict_prefer("filter", "dplyr")
```





```{r init-logging, results="hide", include=FALSE}
flog.appender(appender.file("covid-icu.log"))
flog.threshold(INFO)
flog.info("STARTING script 02: Descriptive Analysis")
flog.info("Logger inited")
```


```{r source-funs}
source(here("src", "01-constants.R"))
source(here("src", "funs", "prepare-data.R"))
source(here("src", "funs", "helper-funs.R"))
source(here("src", "funs", "prop_NA.R"))
source(here("src", "funs", "compare_progression.R"))
source(here("src", "funs", "is_binary.R"))
```


# Import data



**Dataset used: `r params$dataset`**

## Import data

```{r prepare-data, results = "hide", message = FALSE, warning = FALSE}
stopifnot(file.exists(data_processed_path))

d <- read_csv(data_processed_path)
flog.info("Processed data was read.")
flog.info(paste0("Dimensions of data set read are: ", str_c(dim(d), collapse = ", ")))
```


Dimensions of the data set:

```{r}
dim(d)
```




# Exclude "forbidden" columns


These are the "forbidden" variables:

```{r}
forbidden_vars
```

There are `r length(forbidden_vars)` "forbidden" variables.

These variables will be excluded from further analyses.

```{r}
d <-
  d %>% 
  select(-all_of(forbidden_vars))
```


# Compare progression for metric variables

"Forbidden" variables are exlcuded from this and the subsequent analyses.


```{r find-binary-metric-vars}
names_binary_vars <- 
d %>% 
  keep(is_binary, na.rm = TRUE) %>% 
  names()

names_binary_vars

names_metric_vars <-
  d %>% 
  keep(negate(~ is_binary(., na.rm = TRUE))) %>% 
  names()

names_metric_vars
```


## t-Tests sorted by p-value





Name of metric variables (defines as having more than 2 outcome values):

`r names_metric_vars`.


The following table shows the results of t-Tests,
where `verlauf` served as independent variable in each case. 
The dependent variable is one of the metric variables, each in turn.

```{r}
ttest_res <- 
d %>% 
  select(all_of(names_metric_vars), -verlauf) %>% 
  map( ~t.test(.x ~ verlauf, data = d), .id = "var") %>% 
  map_df(tidy) %>% 
  mutate(var = names_metric_vars) %>% 
  mutate(is_signif = p.value < .05) %>% 
  arrange(p.value) 
```


```{r}
ttest_res %>% 
  select(var, estimate, statistic, p.value, is_signif) %>% 
  arrange(-statistic) %>% 
  gt() %>% 
  fmt_number(c(2,3)) %>% 
  fmt_scientific(4)
```


  
  

The sample sizes look quite small though?!

### Check sample size

What's the sample size, counting only complete cases for one particular t-test?

```{r}
d %>% 
  select(id, sat_percent, verlauf) %>% 
  drop_na() %>% 
  nrow()
```

Let's run the t-test manually:

```{r}
d %>% 
  select(id, sat_percent, verlauf) %>% 
  drop_na() %>% 
  t.test(sat_percent ~ verlauf, data = .)
```

The `df` is so small, can that be correct?

Normally, the df is computed by summing the sample sizes of the two groups (minus 2):

$df = n_1 + n_2 - 2$, 

which gives us the complete sample size (minus 2).

Here's a simple test with a well-behaved data set, for testing purposes:

```{r}
t.test(mpg ~ am, data = mtcars)
```

Ah, now I know! The `df` of the Welch-Variant of the t-test is computed differently, yielding a smaller "effective" sample size!

```{r}
t.test(mpg ~ am, data = mtcars, var.equal = TRUE)
```


So, our above computations should be ok!

## Effect size 




```{r}
cohensd_res <- 
d %>% 
  select(all_of(names_metric_vars), -verlauf) %>% 
  map_df( ~ cohens_d(.x ~ verlauf, data = d), .id = "var") %>% 
  arrange(-abs(Cohens_d)) %>% 
  mutate(cles_pr_sup = d_to_cles(Cohens_d)[[1]])
```


```{r}
cohensd_res %>% 
  gt() %>% 
  fmt_number(where(is.numeric))
```




# Comparing nominal variables with regard to progress (bad/good) 

```{r}
cols_to_be_compared_binary <-
  names_binary_vars %>% 
  discard(~ .x %in% c(c("fallnr", "id"), forbidden_vars))
```

There are `r length(cols_to_be_compared_binary)` binary variables to be compared.


```{r}
d_factor <-
  d %>%
  select(verlauf, all_of(cols_to_be_compared_binary)) %>%
  mutate(across(everything(),
                factor))

dim(d_factor)
```




## Chi-Square tests


```{r chi-sq-tests}
chisq_res <- 
  d_factor %>% 
  map2_df(.y = cols_to_be_compared_binary,
       ~ my_chisq3(d = d_factor, x = "verlauf", y = .y),
       .id = "var")  %>% 
  mutate(is_signif = p.value < .05) %>% 
  arrange(p.value) %>% 
  select(-c(method, parameter)) %>% 
  mutate(phi = chisq_to_phi(
    chisq = statistic,
    n = chi_count,
    nrow =chi_nrow,
    ncol = chi_ncol)[[1]],
    cramersv = chisq_to_cramers_v(
      chisq = statistic,
      n = chi_count,
      nrow =chi_nrow,
      ncol = chi_ncol)[[1]])
```


```{r chi-sq-tests-print}
chisq_res %>% 
  select(var, statistic, p.value, is_signif, phi, cramersv) %>% 
  gt() %>% 
  fmt_number(where(is.numeric)) %>% 
  fmt_scientific(p.value)
```




## Interpret effect sizes: Cramers V

```{r add-es-interpret}
chisq_res <- 
chisq_res %>% 
  mutate(cramersv_interpret = interpret_cramers_v(cramersv),
         phi_interpret = interpret_phi(phi))
```


```{r show-es-interpret}
chisq_res %>% 
select(var, cramersv, cramersv_interpret,
       phi, phi_interpret) %>% 
  gt()%>% 
  fmt_number(where(is.numeric))
```


## Significant variables only


```{r}
chisq_res %>% 
  filter(is_signif == TRUE) %>% 
  filter(var != "verlauf") %>% 
   select(var, statistic, p.value, is_signif, phi, cramersv, phi_interpret) %>% 
  gt() %>% 
  fmt_number(where(is.numeric)) %>% 
  fmt_scientific(p.value)
```

